---
title: "language dynamics"
output: html_document
date: "2024-09-01"
---
```{r}
set.seed(1234)
library(igraph)
library(ggplot2)
library(RColorBrewer)
library(rgl)
library(dplyr)
```
##Functions
Function to plot the square lattice
```{r}
plot_grid <- function(S,g,bi=TRUE,label=TRUE) {
  state_colors <- c("red", "black", "white")
  S_colors <- state_colors[S]
  
  layout_matrix <- layout_on_grid(g)  
  
  L <- sqrt(vcount(g))
  node_size <- 230/L  # Adjust this factor to control overall plot size
  
  plot(g, 
       layout = layout_matrix, 
       vertex.size = node_size,               
       vertex.color = S_colors,      
       vertex.shape = "square",
       vertex.frame.color = NA,       
       vertex.label = NA,             
       edge.color = NA,                
       margin = -0.1
  )
  
  if(label) {
  if (bi) { #handle the labels for the two types of models
    legend_labels <- c("Language A", "Language B", "Bilingual")
  } else {
    legend_labels <- c("Language A", "Language B")
    state_colors <- state_colors[1:2]  # Only include the colors for Language A and B
  }
  
  legend("topright", 
         legend=legend_labels,
         col = "black",                      # Border color for the squares in the legend
         pt.bg = state_colors,               # Fill color for the squares in the legend
         pch = 22,                           # Square with border
         pt.cex = 1.5,                       # Size of the squares in the legend
         bty = "n",                          
         cex = 0.8                           
  )
  }
}
```

One step evolution function (bilingual model)
```{r}
step_bi <- function(S,g) {
N <- vcount(g)
#neighbors
n_i <- lapply(1:N,function(v) neighbors(g,v))
n <- sapply(n_i,length)
#fractions of language users
s_A <- sapply(n_i,function(j) (length(which(S[j]==1))))/n
s_B <-sapply(n_i,function(j) (length(which(S[j]==2))))/n
s_AB <- sapply(n_i,function(j) (length(which(S[j]==3))))/n

#We randomly pick a node which has some probability to evolve into a different node
#(has at least a neighbor of a different type)
#idx <- which(s_A!=1 & s_B!=1 & s_AB!=1)
idx_no <- which((s_A==1 & S==1) | (s_B==1 & S==2) | (s_AB==1 & S==3)) #we don't want all neigh of same type as i
idx <- setdiff(1:N, idx_no)
if(length(idx)==0) {
  return(S)
}
i <- sample(idx,1) 
#update probability rule
S[i] <- switch(S[i],
                 sample(c(3,1),size=1,p=c(0.5*s_B[i],1-0.5*s_B[i])),
                 sample(c(3,2),size=1,p=c(0.5*s_A[i],1-0.5*s_A[i])),
                 sample(c(1,2,3),size=1,p=c(0.5*(1-s_B[i]),0.5*(1-s_A[i]),0.5*(s_A[i]+s_B[i])))
)
return(S)
}
```

One step evolution function (agent based Abrams-Strogatz model)
```{r}
step_Str <- function(S,g) {
N <- vcount(g)
#neighbors
n_i <- lapply(1:N,function(v) neighbors(g,v))
n <- sapply(n_i,length)
#fractions of language users
s_A <- sapply(n_i,function(j) (length(which(S[j]==1))))/n
s_B <-sapply(n_i,function(j) (length(which(S[j]==2))))/n

#We randomly pick a node which has some probability to evolve into a different node
#(has at least a neighbor of a different type)
#idx <- which(s_A!=1 & s_B!=1 & s_AB!=1)
idx_no <- which((s_A==1 & S==1) | (s_B==1 & S==2)) #we don't want all neigh of same type as i
idx <- setdiff(1:N, idx_no)
if(length(idx)==0) {
  return(S)
}
i <- sample(idx,1) 
#update probability rule
S[i] <- switch(S[i],
                 sample(c(1,2),size=1,p=c(1-0.5*s_B[i],0.5*s_B[i])),
                 sample(c(1,2),size=1,p=c(0.5*s_A[i],1-0.5*s_A[i]))
               )
return(S)
}
```



##2D regular network
#Bilingual model
```{r}
#side of the network
L <- 25
#number of nodes
N <- L^2

#Number of iterations in the dynamics
T_f <- 5000
```

```{r}
#create the network
g <- make_lattice(dim = 2, length=L, circular = TRUE) #we impose PBC
#states vector (random initialization)
S <- sample(c(1,2,3),N,replace=TRUE) #state 1: first language, 2: second lanugage, 3: both (bilingual)
state_colors <- c("red", "black", "white")  # red for state 1, black for state 2, white for state 3
```


```{r}
plot_grid(S,g)
```





Evolution dynamics
```{r}
#number of time steps
#T <- 10*N^1.8 #N^1.8 scaling of the expected time to reach the extinction of one language
T <- T_f
for (t in 1:T) {
  
  S <- step_bi(S,g)
  if((t %% 1000)==0 | t==1) { #take a shot every 1000 iterations
    png(filename = sprintf("Images/regular_net/bi/regular_net_bi_t=%d.png",t), width = 2000, height = 2000, res = 100)
    plot_grid(S,g,label=FALSE)
    dev.off()
  }
  if(all(S==S[1])) { #If one language dominates
    cat(sprintf("the system reached equilibrium at time %d",t))
    break 
  }
}


plot_grid(S,g)
```
Save plot on file (optional)
```{r}
plot <- TRUE
if(plot) {
png(filename = sprintf("Images/regular_net/bi/regular_net_bi_t=%d.png",t), width = 2000, height = 2000, res = 100)
plot_grid(S,g,label=FALSE)
dev.off()
}
```

###Agent-based version of Abrams-Strogatz
Similar transition rules, but no bilinguism
```{r}
#create the network
g <- make_lattice(dim = 2, length=L, circular = TRUE) #we impose PBC
#states vector (random initialization)
S <- sample(c(1,2),N,replace=TRUE) #state 1: first language, 2: second lanugage
state_colors <- c("red", "black")  # red for state 1, black for state 2
```

```{r}
plot_grid(S,g,bi=FALSE)
```


Evolution dynamics
```{r}
#number of time steps
#T <- 10*N^1.8 #N^1.8 scaling of the expected time to reach the extinction of one language
T <- T_f
for (t in 1:T) {
  
  S <- step_Str(S,g)
  if((t %% 1000)==0 | t==1) {
    png(filename = sprintf("Images/regular_net/Strogatz/regular_net_Str_t=%d.png",t), width = 2000, height = 2000, res = 100)
    plot_grid(S,g,label=FALSE)
    dev.off()
}
  if(all(S==S[1])) { #If one language dominates
    cat(sprintf("the system reached equilibrium at time %d",t))
    break 
  }
}

plot_grid(S,g,bi=FALSE)
```

Save plot on file (optional)
```{r}
if(plot) {
png(filename = sprintf("Images/regular_net/Strogatz/regular_net_Str_t=%d.png",t), width = 2000, height = 2000, res = 100)
plot_grid(S,g,label=FALSE)
dev.off()
}
```






##Small world network
###Bilingual model
```{r}
# Parameters
p <- 0.1    # Rewiring probability
nei <- 1    # Neighborhood radius (each node is connected to 4 nearest neighbors)

#2D small-world network
small_world_net_2d <- sample_smallworld(dim = 2, size = L, nei = nei, p = p, loops = FALSE, multiple = FALSE)

S <- sample(c(1,2,3),N,replace=TRUE)

# Plot the network
plot(small_world_net_2d, 
     layout = layout_on_grid(small_world_net_2d), 
     vertex.size = 5,
     vertex.color=state_colors[S],
     vertex.label = NA,
     edge.color = "grey",
     edge.width = 0.5)

plot_grid(S,g)
```
```{r}
if(plot) {
png(filename = sprintf("Images/smallw/bi/smallw_bi_t2=%d.png",t), width = 2000, height = 2000,res=100)
plot_grid(S,g,label=FALSE)
dev.off()
}
```

Evolution dynamics
```{r}
#number of time steps
T <- T_f
for (t in 1:T) {
  
  S <- step_bi(S,g)
  if((t %% 1000)==0 | t==1) { #take a shot every 1000 iterations
    png(filename = sprintf("Images/smallw/bi/smallw_bi_t=%d.png",t), width = 2000, height = 2000, res = 100)
    plot_grid(S,g,label=FALSE)
    dev.off()
  }
  if(all(S==S[1])) { #If one language dominates
    cat(sprintf("the system reached equilibrium at time %d",t))
    break 
  }
}


plot_grid(S,g)
```

```{r}
if(plot) {
png(filename = sprintf("Images/smallw/bi/smallw_bi_t=%d.png",t), width = 2000, height = 2000, res = 100)
plot_grid(S,g,label=FALSE)
dev.off()
}
```

################
```{r}
plot(small_world_net_2d, 
     layout = layout_on_grid(small_world_net_2d), 
     vertex.size = 5,
     vertex.label = NA,
     edge.color = "grey",
     edge.width = 0.5,
     vertex.color=state_colors[S])
```
#####################




###Abrams-Strogatz
```{r}
# Parameters
p <- 0.1    # Rewiring probability
nei <- 1    # Neighborhood radius (each node is connected to 4 nearest neighbors)

#2D small-world network
small_world_net_2d <- sample_smallworld(dim = 2, size = L, nei = nei, p = p, loops = FALSE, multiple = FALSE)

S <- sample(c(1,2),N,replace=TRUE)

# Plot the network
plot(small_world_net_2d, 
     layout = layout_on_grid(small_world_net_2d), 
     vertex.size = 5,
     vertex.color = state_colors[S],
     vertex.label = NA,
     edge.color = "grey",
     edge.width = 0.5)

plot_grid(S,g,bi=FALSE)
```
```{r}
if(plot) {
png(filename = sprintf("Images/smallw/Strogatz/smallw_Strogatz_t=%d.png",t), width = 2000, height = 2000, res = 100)
plot_grid(S,g,label=FALSE)
dev.off()
}
```

Evolution dynamics
```{r}
#number of time steps
T <- T_f
for (t in 1:T) {
  S <- step_Str(S,g)
  if((t %% 1000)==0 | t==1) { #take a shot every 1000 iterations
    png(filename = sprintf("Images/smallw/Strogatz/smallw_Strogatz_t=%d.png",t), width = 2000, height = 2000, res = 100)
    plot_grid(S,g,label=FALSE)
    dev.off()
  }
  if(all(S==S[1])) { #If one language dominates
    cat(sprintf("the system reached equilibrium at time %d",t))
    break 
  }
}

plot_grid(S,g,bi=FALSE)
```




```{r}
if(plot) {
png(filename = sprintf("Images/smallw/Strogatz/smallw_Strogatz_t=%d.png",t), width = 2000, height = 2000, res = 100)
plot_grid(S,g,label=FALSE)
dev.off()
}
```
